% Section 9 --- Conclusion
% normalisation needed (no smart-quotes, no underscores, no new citations).
% The final-sentence general-validation phrasing is the allowed negated form.
\section{Conclusion}\label{sec:conclusion}

widom-atlas began from a simple problem: a Widom insertion number is easy to compute, but hard to trust unless the calculation that produced it is completely specified. The Henry constant and zero-coverage isosteric heat are not merely outputs of an estimator. They are outputs of a structure, an atom typing, an energy model, a gas model, a mixing rule, an electrostatics convention, a backend, a sampling procedure, and a reference observable. The central claim of this paper is that this recipe must be treated as the scientific object.

The v0.4 audit gives a narrow result, presented as method-plus-audited-examples rather than a fifteen-branch certification. Two of fifteen verdict-affecting branches pass the strict criterion: the all-silica MFI + Ar positive control and the UiO-66 + CO\(_2\) branch under genuine UFF. Five of fifteen pass the Tier-B physical-accuracy criterion, adding HKUST-1 + CO\(_2\) under genuine UFF, the Si-CHA + CO\(_2\) Bai 2013 control, and the UiO-66 + CO\(_2\) Maia 2023 explicit-hydrogen variant. These results show that widom-atlas can reproduce experiment when the recipe and the reference are compatible --- and twice that conclusion survived only because the audit caught a defect first. The MFI + Ar control passed the strict band even as originally locked, but only as a fortuitous cancellation of four compounding errors; the cross-engine audit of Section~\ref{sec:v06} caught the silicon double-count and rebuilt the recipe clean (Finding~F1). The same audit caught a generator silently mislabelling the HKUST-1 and UiO-66 framework force fields, whose correction to genuine UFF moved both from apparent open-metal-site failures to physical-accuracy passes (one strict). They do not show that Widom insertion with standard force fields is generally predictive.

The negative branches are not failures of the same kind. The remaining open-metal-site systems --- the Mg-MOF-74 branches --- expose genuine force-field transferability limits. Some zeolite branches expose reference or heat-of-adsorption sensitivity. The cationic-zeolite targets remain incomplete recipes because the cation-containing structures and cation models are not yet locked. Na-Rho is method-blocked because a rigid closed-state Widom calculation and a trapdoor adsorption experiment are not the same statistical-mechanical observable. Treating all of these cases as ordinary failures would erase the physics that the audit is designed to reveal.

The useful product of widom-atlas is therefore not a larger table of adsorption numbers. It is a governed classification of those numbers. A branch can be a strict validation point, a weaker physical-accuracy point, a force-field disagreement, a reference-observable mismatch, a structural blocker, or an ensemble/method mismatch. This classification tells a screening workflow whether a score is a calibrated observable, a useful but weaker physical comparison, or a number that should not be trusted for the claimed purpose.

The present atlas is not finished. Flexible trapdoor zeolites require methods that sample the relevant framework states. Cationic zeolites require locked cation structures and cation force fields. Open-metal-site MOFs require energy models capable of the missing binding physics. Consensus references are needed where the literature observable itself is uncertain. Machine-learned potentials may become useful, but only as new energy-model recipes that must be validated on their own terms; Section~\ref{sec:v06} takes a first step in exactly that spirit, on the engine and the potential that CuspAI ships. None of these future tasks changes the v0.4 verdict.

The conclusion is therefore deliberately modest. widom-atlas is not scientifically validated as a general predictive adsorption tool. It is a reproducible way to decide whether a particular Widom adsorption number is a trustworthy zero-coverage observable or merely the output of an untested recipe. That distinction is the result.
