% Appendix E --- v0.5 non-load-bearing notes. Populated 2026-06-09. Keeps v0.5
% clearly future-work and non-load-bearing; does not make any v0.5 finding
% load-bearing on the v0.4 verdict.
\section{v0.5: non-load-bearing extension notes}\label{app:v05}

\textbf{Version map.} v0.4 is the locked classical audit matrix; v0.4.2 is the corrected rebuild and provenance layer for the audited examples (the genuine-UFF 2a and 3a, the cross-engine-checked 4c, and the clean 6c rebuild); v0.5 contains non-load-bearing extension notes only (this appendix); v0.6 is the kUPS and MLIP governance companion, summarised in Section~\ref{sec:v06}. Only the stated locked branches and committed evidence affect the v0.4/v0.4.2 verdicts.

The work described in this appendix is future work, and it does not change the v0.4 verdict. Any v0.5 feature or review branches that exist are non-load-bearing: none of them is used in, or alters, the v0.4 result reported in the main text.

The future extensions under consideration are: a polarisable treatment of Mg-MOF-74 + CO\textsubscript{2}; a quantitative multi-source consensus-reference layer; completion of cation-containing zeolite structures and cation models; and machine-learned-potential or ASE-calculator routes. Each is a future extension only, and each would have to be validated as a new recipe on its own terms before it could affect any verdict.

None of these extensions changes the v0.4 disposition. None changes the strict-pass result of two of fifteen branches. None changes the Tier-B physical-accuracy result of five of fifteen branches. None converts the method-blocked Na-Rho (5b) scalar comparison into a pass or a fail. And none makes the reference-audited 5c cationic-zeolite targets a completed validation family.
