\section{Thesis}\label{thesis}

A published force-field parameter table is \textbf{not} a complete
simulation recipe. The load-bearing ingredients are frequently
unpublished: a short-range protection convention, a mixing rule applied
to the framework atoms, a dispersion correction, a blocking radius, an
energy-reference convention. An ungoverned pipeline that guesses any of
them wrong returns a confident, wrong number with no warning. This is
the \textbf{domain-of-validity / realization-gap} failure class.
The v0.4/v0.4.2 classical audit established this failure class on locked
classical force-field recipes; \textbf{v0.6 retargets the same governance
logic onto CuspAI's own engine (kUPS) and the machine-learning
interatomic potential (MLIP) it ships} --- and shows a governance
layer that \emph{characterizes, discriminates, and where the evidence
supports it, screen-passes a governed remediation.}

\textbf{Reader note on labels.} Branch codes such as \texttt{6c} and
\texttt{4c} are internal locked-recipe identifiers, not chemical formulae
or universal benchmark names: each fixes one complete Widom-insertion
recipe (structure, guest model, force field, charges, mixing rule,
electrostatics, backend, sampling, and reference). Branch 6c is the
all-silica MFI + Ar recipe used as the positive control; branch 4c is the
Si-CHA + CO\textsubscript{2} charged-zeolite control. ``Native'' is this
project's own independent Python Widom estimator (not a generic term),
reported alongside the external Monte Carlo engine RASPA3. The version
labels are internal repository milestones: v0.4 and v0.4.2 are the locked
and corrected classical-audit layers, v0.5 is non-load-bearing
consolidation or extension material, and v0.6 is this kUPS and MLIP
governance extension.

Three results carry the report: 1. \textbf{A2/F1} --- the originally-locked positive
control \emph{double-counted silicon}; its strict-band agreement was a fortuitous four-error
cancellation; rebuilt clean (single-source, Talu's calibration Olson
geometry) it reproduces the corrected reference \(K_H = 0.200\)
cross-engine (native 0.200, RASPA3 0.199) (Finding~F1). 2. \textbf{A2-F2} --- to our knowledge, the first published
independent validation of kUPS's (six-week-old, unreleased) Widom
estimator. 3. \textbf{WPB} --- a four-configuration MLIP demonstration:
discriminate two failure modes, screen-pass a dispersion-corrected
route, and show the failure generalizes across MLIP families.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{Cross-engine Widom parity
(WPA-A2)}\label{cross-engine-widom-parity-wpa-a2}

\subsection{System and recipe}\label{system-and-recipe}

6c positive control: \textbf{MFI (all-silica) + Ar}, García-Pérez 2007
framework + Talu-Myers 2001 cross-pair. The complete locked recipe
(recovered verbatim from the RASPA3 evidence
\texttt{force\_field.json}):

{%
\begin{longtable}[]{@{}lll@{}}
\toprule\noalign{}
element & self LJ (ε/K, σ/Å) & source \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Ar & 119.8 / 3.405 & Talu-Myers 2001 \\
O & 53.0 / 3.3 & TraPPE-zeo, Bai 2013 \\
Si & 22.0 / 2.3 & TraPPE-zeo, Bai 2013 \\
\end{longtable}
}

Ar--O is an \textbf{explicit} binary (93.0 K / 3.335 Å, Talu-Myers);
\textbf{MixingRule Lorentz-Berthelot}, \textbf{truncation shifted},
cutoff 12 Å, \textbf{TailCorrections false}, no electrostatics.
Critically, the Lorentz-Berthelot rule applied to the framework
self-parameters generates a \textbf{second} interaction ---
\textbf{Ar--Si (51.34 K / 2.8525 Å)} --- in addition to the explicit
Ar--O term. This is the \emph{as-locked} recipe, recorded verbatim;
Finding~F1 (§2.4) shows this Ar--Si term is a \emph{silicon double-count}
(Talu's Ar--O already folds silicon's dispersion into oxygen) and gives
the clean single-source rebuild that reproduces the corrected reference.

\subsection{The clean metric}\label{the-clean-metric}

Engines report different units (kUPS: K\_H in Å³/eV; RASPA3/native:
mol·kg⁻¹·bar⁻¹), so the comparison is on the engine-agnostic
\textbf{⟨exp(−βU)⟩}, with K\_H and q\_st reported in common units.

\subsection{Result}\label{result}

{%
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Engine (full recipe unless noted)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
⟨W⟩
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
K\_H (mol·kg⁻¹·bar⁻¹)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
q\_st (kJ/mol)
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
RASPA3 (locked reference) & --- & \textbf{0.2070} & \textbf{16.03} \\
native independent (Ar--O \textbf{+ Ar--Si}) & 9.54 & \textbf{0.2093}
(+1.1 \%) & 15.91 (−0.12) \\
kUPS Widom (Ar--O \textbf{+ Ar--Si}) & 11.08 & 0.2432 (+16.2 \%) & 16.25
(+0.22) \\
native (Ar--O \textbf{only}, naive) & 5.24 & 0.1149 (\textbf{1.8× low})
& 14.36 \\
kUPS (Ar--O only) & 5.98 & 0.1313 & --- \\
\end{longtable}
}

{\footnotesize\noindent Percent differences are relative to the stated comparator. The native
\(K_H\) offset (\(+1.1\,\%\)) is relative to the RASPA3 locked reference; the kUPS \(K_H\) offset
(\(+16.2\,\%\)) is quoted relative to the convention-matched native shifted estimator when discussing
the shifted-versus-truncated convention gap (it is \(+17.5\,\%\) relative to RASPA3). The \(q_{st}\)
offsets in parentheses are relative to RASPA3.\par}

\subsection{Finding F1 --- a silicon double-count caught by audit (and a
clean rebuild)}\label{finding-f1-the-arsi-term-is-load-bearing-1.82-on-k_h}

The originally-locked 6c recipe \textbf{double-counted silicon}, and
agreed with its reference only as a fortuitous cancellation of four
compounding errors. Talu--Myers' Ar--O cross-parameter (93.0 K) is the
Lorentz--Berthelot combination of argon with an oxygen-only
\emph{effective} oxygen (72.2 K, Talu Table 1) that already folds
silicon's dispersion into oxygen --- Talu's model is oxygen-only by
construction (Talu \S2: ``ignores partially concealed silicon atoms and
lumps their dispersion energy with the oxygen atoms''). The locked recipe
nonetheless added an explicit Ar--Si from an active TraPPE-zeo framework
silicon, counting silicon twice (\(\times 1.82\) on \(K_H\)). That
spurious term compensated three further errors --- an idealized (IZA)
rather than the calibration (Olson) geometry (\(\sim\!-25\,\%\)), an
under-converged 12~\AA{} tail-off cutoff (\(\sim\!-25\,\%\) more), and a
reference value mis-derived by \(\sim\!12\,\%\) (0.224 vs the correct
0.200) --- which multiplied back to roughly the (wrong) reference.
Removing all four and running the coherent \textbf{single-source} recipe
(pure Talu, oxygen-only, no Ar--Si) at a converged cutoff on Talu's own
(Olson) geometry reproduces the \textbf{corrected} reference
(\(K_H = 0.200\)) exactly and \textbf{cross-engine}: native 0.200,
RASPA3 0.199 (parity 0.5\,\%), \(q_{st}\) 14.97 (within 0.7 kJ/mol of
calorimetry). In strict-band terms the as-locked hybrid lands in-band against \emph{both} the
mis-derived 0.224 and the corrected 0.200 reference; the rebuild changes the recipe's coherence, not the
pass/fail verdict.

{%
\begin{longtable}[]{@{}lll@{}}
\toprule\noalign{}
6c recipe / setup & \(K_H\) (mol/kg/bar) & note \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
locked hybrid (IZA, 12~\AA{} tail-off) & 0.207 & in-band by a fortuitous cancellation (also passes 0.200) \\
\(-\) double-counted silicon & 0.115 & pure Talu, oxygen-only (IZA, 12~\AA{}) \\
\(+\) converged 24~\AA{} + tail & 0.154 & numerical defect fixed (IZA) \\
\(+\) Olson calibration geometry & \textbf{0.200} & native; RASPA3 0.199 (parity 0.5\,\%) \\
\textbf{corrected reference} (Dunne 1996) & \textbf{0.200} & strict window [0.159, 0.252] \\
structure sensitivity (single-source) & 0.200 / 0.205 / 0.154 & Olson / van Koningsveld / IZA \\
\end{longtable}
}

(The locked hybrid evidence is retained intact as the documented
negative case.)

\subsection{Finding F2 --- independent cross-engine validation of
kUPS's Widom estimator}\label{finding-f2-the-first-published-independent-validation-of-kupss-widom-estimator}

kUPS ships a Widom test-particle module (\texttt{kups.mcmc.widom}; in
source \texttt{main}, absent from the released PyPI
\texttt{kups\ 1.0.1}; no published external validation). v0.6 is, to our
knowledge, the first \emph{published} independent cross-engine check of
this specific module --- distinct from standalone Widom packages (e.g.\ the
\texttt{widom} ASE add-on) and from machine-learned-potential adsorption
benchmarks such as ODAC25 (Sriram et al.\ 2025, arXiv:2508.03162), based on public-source
searches (the kUPS source carries internal unit tests only; the open
tracker was inspected 2026-06-12 --- one unrelated feature issue).

\begin{itemize}
\item
  \textbf{The estimator is correct.} \texttt{widom\_test()} returns lnα
  = −βΔU for a ghost insertion ⇒ \textbf{W ≡ exp(−βU)} --- identical to
  the textbook Widom weight; the μ\_ex / K\_H / q\_st reductions match
  widom-atlas's definitions. Verified by reading the source \emph{and}
  by the parity run.
\item
  \textbf{q\_st is validated to 0.22 kJ/mol} (16.25 vs 16.03 RASPA3;
  15.91 native). The energies are right.
\item
  \textbf{The +16 \% K\_H is a shifted-vs-truncated LJ convention ---
  root-caused, not a normalization bug.} kUPS's
  \texttt{lennard\_jones\_energy} is \textbf{truncated, not
  energy-shifted} (\texttt{4ε(c6²−c6)} masked at cutoff, no
  \texttt{−U(rc)} term); our config (\texttt{tail\_correction:\ false},
  no \texttt{truncation\_radius}) selects that plain variant. RASPA3's
  recipe is \texttt{TruncationMethod:\ shifted}, and the native
  estimator shifts too. The unshifted potential is more attractive by
  U(rc) per interaction. Toggling the shift off in native reproduces
  kUPS:

  {%
  \begin{longtable}[]{@{}llll@{}}
  \toprule\noalign{}
  native (with-Si) & ⟨W⟩ & K\_H & q\_st (kJ/mol) \\
  \midrule\noalign{}
  \endhead
  \bottomrule\noalign{}
  \endlastfoot
  shifted (RASPA convention) & 9.54 & 0.2093 & 15.91 \\
  truncated / no-shift (kUPS convention) & 11.35 & 0.2490 & 16.34 \\
  kUPS Widom & 11.08 & 0.2432 & 16.25 \\
  \end{longtable}
  }

  The implied constant shift δ = RT·ln(11.35/9.54) = \textbf{0.43
  kJ/mol} (predicted ≈0.35) explains \textbf{both} the +16 \% K\_H
  \textbf{and} the +0.34 kJ/mol q\_st: matching kUPS's convention,
  native-no-shift agrees with kUPS to \textbf{+2.4 \% on K\_H (within
  the 3.4 \% SEM) and 0.09 kJ/mol on q\_st}. Neither engine is wrong ---
  but the shift convention is a load-bearing recipe element worth
  \textasciitilde16 \% on K\_H, which must be locked and matched. The
  exponential sensitivity of K\_H to a constant energy offset, and the
  resulting need to lock the tail/truncation convention, are established
  (Jablonka, Ongari \& Smit 2019, \emph{J. Chem. Theory Comput.}\ \textbf{15},
  5635; Dubbeldam et al.\ 2016, \emph{Mol. Simul.}\ \textbf{42}, 81); what is
  new here is the cross-engine \emph{quantification} (\(\delta \approx 0.43\)
  kJ/mol). Does
  \textbf{not} affect q\_st-as-physics or the OOD governance demo
  (energy path).
\end{itemize}

The recipe all three engines share for this parity is the as-locked
hybrid now reframed in F1 as a silicon double-count: F2 establishes that
independent estimators agree on a \emph{common} recipe (to \(1.1\,\%\) on
\(K_H\)), not that the recipe reproduces experiment --- that correction
lives entirely in F1, and the two findings are consistent.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{Blocking-sphere governance
(WPA-A3)}\label{blocking-sphere-governance-wpa-a3}

kUPS exposes the v0.5 short-range-protection lesson as an explicit API
--- \texttt{blocking\_spheres} (per-host hard-sphere exclusions). So the
convention is no longer a claim of ours; it is CuspAI's own engineering.
A sweep on the 6c system (64 spheres on framework-O positions):

{%
\begin{longtable}[]{@{}lllll@{}}
\toprule\noalign{}
blocking & radius (Å) & K\_H (Å³/eV) & ΔK\_H & q\_st (kJ/mol) \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
off & --- & 1.798×10⁷ & --- & 16.25 \\
on & 2.0 & 1.856×10⁷ & +3.2 \% & 16.34 \\
on & 3.0 & 1.812×10⁷ & +0.8 \% & 16.28 \\
\end{longtable}
}

\textbf{Finding.} On a physically-repulsive LJ recipe blocking is a
\textbf{no-op} (ΔK\_H \textless{} SEM, q\_st unchanged): the potential
already excludes overlaps, so a hard sphere on top adds nothing. The
corollary is the governance lesson --- \textbf{blocking is load-bearing
exactly when the recipe has an \emph{unphysical} short-range attraction
the potential does not repel} (open-metal-site over-binding; the v0.5 1b
exp-6 case). A recipe's blocking configuration is therefore part of the
recipe and must be locked: invisible on an all-silica LJ host, decisive
on an OMS host. The small +0.8\ldots+3.2 \% K\_H nudge from blocking is
an insertion-count effect (a distinct, much smaller thing than F2's +16
\%, which is the LJ shift convention), within the SEM and not affecting
q\_st.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{MLIP out-of-distribution governance (WPB) --- the
headline}\label{mlip-out-of-distribution-governance-wpb-the-headline}

\subsection{Setup}\label{setup}

Widom insertions of \textbf{Ar into all-silica CHA}, N = 120, seed 0, T
= 298.15 K. A classical UFF LJ baseline scores the \textbf{same}
insertion geometries as each MLIP, so all differences are the potential,
not the sampling. Two transparent OOD flags (hard-overlap at \textless{}
0.80 σ\_min = 2.762 Å; energetic-anomaly at U \textless{} −25 kJ/mol or
U \textless{} U\_classical − 50) and a \textbf{pre-registered verdict
schema} with two refusal classes: \textbf{Class I} (flagged
Boltzmann-weight fraction ≥ 0.5 → OOD over-binding) and \textbf{Class
II} (min U \textgreater{} 0 or q\_st ≤ RT where physisorption is
established → under-binding). The schema was written before the results
so refusal is rule-based, not retrofit.

The governance premise is the model authors' own statement: CuspAI's
\texttt{kUPS-mace-jax} model card declares MACE-MPA-0 \emph{``not
trained for isolated molecules''} --- scoring an adsorbate is
out-of-domain by the authors' own words. What is refused is
\textbf{ungoverned usage of a configuration}, never a model.

\subsection{Result --- four configurations, discriminate then
remediate}\label{result-four-configurations-discriminate-then-remediate}

{%
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Configuration
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
min U (kJ/mol)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
flagged-weight
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Verdict
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
MACE-MP-small (2023) & −19.3 & 0.995 & \textbf{REFUSE --- Class I}
(over-binds overlaps) \\
MACE-MPA-0 bare (kUPS's shipped export) & +8.5 & 0.0 & \textbf{REFUSE
--- Class II} (no physisorption) \\
MACE-MPA-0 \textbf{+ D3(BJ)} & −5.03 & 0.0001 & \textbf{screen-pass}
(physical but shallow well; accuracy unassessed) \\
UMA \texttt{uma-s-1.1}, task = omat (fairchem flagship) & −292.8 & 1.0 &
\textbf{REFUSE --- Class I} (erratic OOD) \\
\end{longtable}
}

Paired energies at identical \emph{open} (\textasciitilde4 Å, no
overlap) insertions make the mechanism airtight:

{%
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1429}}
  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1429}}
  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1429}}
  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1429}}
  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1429}}
  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1429}}
  >{\raggedright\arraybackslash}p{(\linewidth - 12\tabcolsep) * \real{0.1429}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
insertion
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
min\_dist Å
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
classical
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
small (2023)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
MPA-0 bare
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
+ D3(BJ)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
UMA omat
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\#76 & 4.33 & −16.2 & −0.3 & +10.7 & +0.2 & −58.3 \\
\#95 & 4.23 & −13.4 & −0.1 & +8.5 & −1.4 & −39.3 \\
\#51 & 3.99 & −17.8 & −0.4 & +10.5 & −1.0 & +121.3 \\
\end{longtable}
}

\subsection{Interpretation}\label{interpretation}

\begin{itemize}
\item
  \textbf{MACE-MP-small (2023)} hallucinates \emph{attractive} energies
  inside the repulsive overlap region --- its global minimum (−19.3)
  sits \emph{at} a hard overlap, and the flagged insertions carry
  \textbf{99.5 \%} of the partition weight. At physical sites its
  attraction is ≈ 0 (−0.1\ldots−0.4): the ``binding'' is a pure overlap
  artifact, not physisorption. \textbf{Class I.}
\item
  \textbf{MACE-MPA-0 bare (2024, kUPS's shipped weights)} \emph{fixed}
  the overlap hallucination --- it is correctly repulsive at overlaps
  --- but is repulsive \emph{everywhere} (min +8.5), finding no
  physisorption well. This is the bare-PBE signature: no −C₆/r⁶
  dispersion tail. \textbf{Class II.}
\item
  \textbf{MACE-MPA-0 + D3(BJ)} --- pairing the \emph{same} weights with
  the standard D3 Becke-Johnson dispersion correction (xc = PBE, cutoff
  40 Bohr, torch-dftd 0.5.3) restores a physical well (min −5.03) at a
  non-overlap distance (3.37 Å; flagged-weight 0.0001). \textbf{Screen-passes; accuracy unassessed.} The
  governance lesson crystallizes: \emph{dispersion-correction on/off is
  a load-bearing recipe element, and ungoverned pipelines silently
  differ on it.} kUPS's potential list contains no dispersion term --- a
  finding, recorded, not a criticism.
\item
  \textbf{UMA \texttt{uma-s-1.1} (deployment-class flagship)} ---
  fairchem's multi-task foundation model (ODAC23 is one of its five
  training sets, so it is the UMA \emph{and} ODAC leg).
  \texttt{task\_name} is a load-bearing recipe element (five DFT levels
  of theory), and it produces a clean \textbf{task × system matrix}:

  {%
  \begin{longtable}[]{@{}
    >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
    >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
    >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
    >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
    >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}@{}}
  \toprule\noalign{}
  \begin{minipage}[b]{\linewidth}\raggedright
  task
  \end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
  system
  \end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
  Gate 1
  \end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
  min U
  \end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
  Verdict
  \end{minipage} \\
  \midrule\noalign{}
  \endhead
  \bottomrule\noalign{}
  \endlastfoot
  omat & Ar/CHA & −0.19 ✅ & −292.8 & REFUSE --- Class I (erratic OOD
  overlaps) \\
  odac & Ar/CHA & −2.73 ❌ & --- & WITHHELD (inconsistent at equivalent
  sites) \\
  \textbf{odac} & \textbf{CO₂/CHA} & +0.06 ✅ & \textbf{−15.66} &
  \textbf{screen-pass} (in-domain, N=60 small sample; q\_st 17.7 vs
  anchor 21.0, −3.3 below strict) \\
  omat & CO₂/CHA & +0.04 ✅ & +1.16 & REFUSE --- Class II (no CO₂
  well) \\
  \end{longtable}
  }

  The \textbf{odac} head (trained on MOF/CO₂/H₂O direct-air-capture
  data) finds a \textbf{physical CO₂ well (−15.7 kJ/mol, screen q\_st
  17.7; experimental CO₂/CHA ≈ 20--30) on CO₂/CHA --- CO₂ is
  adsorbate-in-domain, though the all-silica CHA host is outside the MOF host class it was trained on → governance screen-passes it}. The screen q\_st is 17.7 kJ/mol against a reference anchor of ≈21 (≈3 kJ/mol low, −3.3 below strict): a screen-pass, not an accuracy certificate. On Ar/CHA (a noble gas it
  never trained on) the \emph{same} head is internally inconsistent ---
  2.7 kJ between symmetry-equivalent open sites --- which Gate 1 catches
  → \textbf{withheld}. The \textbf{omat} (materials) head under-binds
  adsorbates (no CO₂ well; erratic deep-overlap hallucinations on Ar).
  \textbf{The model passes exactly where its documented training domain
  says it should, and is refused/withheld everywhere else} --- the
  governance respects the model card rather than fighting it.
  Separately, the OOD-overlap failure on Ar is not MACE-specific: it
  generalizes across MLIP families (MACE-MP-small and UMA-omat alike).
\end{itemize}

\subsection{Validity gates (why the numbers can be
trusted)}\label{validity-gates-why-the-numbers-can-be-trusted}

\begin{itemize}
\tightlist
\item
  \textbf{MACE Gate 1 (reference offset).} A +8.5 floor could be a
  broken differencing. Ruled out: Ar is in the model (Z = 18, 89-element
  table); \textbf{U(Ar isolated, 12.9 Å) = 0.000 kJ/mol} ---
  differencing exact. The +floor is real missing-dispersion physics.
\item
  \textbf{UMA --- two hazards, both caught by Gate 1 before any
  verdict.} (1) UMA conditions on a \textbf{per-graph} global
  charge/spin/task embedding, so naive three-graph differencing does not
  cancel --- a task-dependent far-field offset (omat +474, odac −245,
  omol +119 kJ at 12.9 Å). Fixed with a \textbf{same-graph reference}
  U(r) = E(host+Ar@r) − E(host+Ar@ref). (2) A non-periodic cut cluster
  is itself OOD for a periodic materials model (+800\ldots+2900 kJ at
  open sites) despite passing the narrow far-field check; fixed by
  \textbf{periodic} in-domain evaluation with an open-pore reference
  (Gate 1 = −0.19 kJ; a 0.09 Å overlap → +21 572 kJ, correctly
  repulsive). \textbf{task\_name is a locked, load-bearing recipe
  element} (UMA carries five DFT levels of theory; we lock
  \texttt{omat}). The governance harness thus validates \emph{itself}
  for an engine before it is permitted to judge that engine.
\end{itemize}

\subsection{Provenance (pinned)}\label{provenance-pinned}

MACE-MPA-0 checkpoint SHA-256{[}:16{]} \texttt{75428afe3a1d7d80}; kUPS
export \texttt{mace-mpa-0-medium\_32.zip} SHA-256{[}:16{]}
\texttt{c5eb645f2dc2c904}; D3 = D3(BJ)/PBE/40 Bohr via torch-dftd 0.5.3;
UMA = \texttt{uma-s-1.1} (extensivity-safe, not \texttt{uma-s-1}),
fairchem-core 2.21.0. N, seed, T, CIF in every summary JSON.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{Methods honesty}\label{methods-honesty}

\begin{itemize}
\tightlist
\item
  \textbf{Numerical guards are recipe elements.} Two self-corrections
  are documented and retained: a spurious −50 weight-clip that
  manufactured a fake Class-I verdict for bare MPA-0 (discarded), and an
  eV/kJ unit mismatch in the Boltzmann weight that inflated exponents
  ×96 and tripped the exp(700) overflow cap spuriously (fixed with RT =
  2.478 kJ/mol; after the fix no leg hits the cap --- confirming the
  cap-firing was the bug, not deep physics). Verdict \emph{directions}
  were unchanged; the quantitative weights are now physical.
\item
  \textbf{Convergence.} kUPS: 25 000 ghost insertions, 12 blocks, K\_H
  SEM ≈ 3.4 \%; the +16 \% is a systematic. native: 1--2×10⁶ insertions,
  ⟨W⟩ \textless{} 0.5 \%. The analytical LJ tail was tested on native
  (moves K\_H only +8 \%), confirming the recipe's
  \texttt{TailCorrections:\ false} and ruling tail out as the parity gap
  driver --- the Ar--Si term, not tail, was the 1.8×.
\item
  \textbf{Native charged-Ewald limitation (classical matrix).} On the
  Si-CHA + CO\(_2\) charged control (4c) the native estimator and RASPA3
  agree to \(\sim\!9\%\) on K\_H and \(\sim\!1\%\) on q\_st (inside the
  experimental window); on the LJ-only MFI + Ar control they agree to
  0.5\%. On HKUST-1, by contrast, the native charged-Ewald path shows a
  large empirical discrepancy against RASPA3 --- about a factor of four on
  the identical recipe. A suspected non-orthogonal minimum-image
  explanation was tested: at the actual supercells and cutoffs of the
  published runs, a brute-force image check found no within-cutoff
  mis-imaged pairs for HKUST-1, Si-CHA, or MFI, so the root cause is not
  established. The classical MOF branches are therefore reported from
  RASPA3 (2a, 3a) or RASPA2 (1a), or presented as audited examples with
  this limitation stated (1b, 1c, 1d, 2b, 3b); 4c is retained because it is
  directly cross-engine checked, not because of any orthogonality
  assumption. For
  4c itself the strict-vs-Tier-B verdict is engine-marginal: the q\_st
  deviation straddles the \(\pm2.0\) boundary (native \(+1.99\), inside;
  RASPA3 \(+2.24\), outside), so we report it as Tier-B --- a worked
  example, on a charged system, of the verdict fragility this harness
  exists to expose. No result in this report depends on the native
  charged-MOF path.
\item
  \textbf{A third silent-substitution defect (1c).} Beyond the 6c silicon
  double-count (F1) and the HKUST-1/UiO-66 generator mislabel (framework
  atoms stamped ``UFF'' but carrying guest/zeolite parameters; corrected to
  genuine UFF, both MOFs reproduce experiment), the audit also caught
  branch 1c (Becker Mg-MOF-74): its as-run framework LJ was plain UFF with
  a placeholder Mg \(\epsilon = 5.0\) K, not the ``reduced'' Becker values
  its label claimed. Retained as a FAIL example, not rebuilt --- the same
  defect class the harness exists to expose.
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{Open scientific questions}\label{open-scientific-questions}

\begin{itemize}
\tightlist
\item
  \textbf{kUPS K\_H +16 \% --- RESOLVED (shift convention).} Root-caused
  to kUPS's truncated (unshifted) LJ vs RASPA/native shifted; δ ≈ 0.43
  kJ/mol reconciles K\_H to 2.4 \% and q\_st to 0.09 kJ/mol (§2.5). Open
  only as an upstream documentation point: kUPS's default LJ is
  unshifted.
\item
  \textbf{Ewald leg (4c, Si-CHA + CO₂) --- DONE, read via the convention
  matrix.} TraPPE-CO₂ + framework charges + Ewald, 14 Å + tail.
  \textbf{q\_st validated to 0.01 kJ/mol} (22.98 vs 22.99). For K\_H, a
  fresh same-session native re-run (the 2.223 was the v0.4 \emph{lock};
  the re-run gives 2.215, reproducing it to 0.3 \%) toggles the
  conventions: \textbf{shift = +11.6 \%, tail = +11.9 \%} (the shift is
  \emph{not} negligible at 14 Å --- earlier-draft correction). kUPS uses
  a truncated LJ and \textbf{does apply a tail} (tested:
  \texttt{tail\_correction:\ false} drops K\_H by \(14\,\%\) (\(2.271 \to 1.952\)), q\_st by \(0.49\) --- a
  ``no-tail-in-Widom'' hypothesis is \emph{falsified}). So its +2.2 \%
  vs the shifted lock is a \textbf{cancellation}: read on the same
  (truncated+tail) convention, \textbf{kUPS is −8.1 \% vs native} (−11.7
  \% with tail also off). That residual is \textbf{q\_st-invariant
  (Δq\_st ≤ 0.3 kJ/mol) → a normalization-flavoured ⟨W⟩ scale}
  (volume/insertion-count), 4c-specific and larger than 6c's −2.4 \%;
  contributors include a tail-implementation difference (kUPS +16.3 \%
  vs native +11.9 \%), the Ewald convention, and CO₂ orientation
  (\texttt{native\_4c\_convention\_matrix.json}).
\item
  \textbf{OMS blocking positive case.} The dramatic blocking
  demonstration (collapsing a divergent open-metal K\_H to a physical
  value) needs a kUPS-representable OMS recipe with charges (1c Becker,
  or 2a UFF Cu + EPM2); the exotic 1a/1b/1d Buckingham forms are not
  representable in kUPS (LJ/Coulomb/Morse/MACE only).
\item
  \textbf{WPB through kUPS's in-engine JAX MACE --- DONE.} Loaded kUPS's
  \texttt{TojaxedMliap} export and scored identical geometries; agrees
  with the torch path to \textbf{max rel 2.1×10⁻⁷}
  (\texttt{jax\_torch\_parity.json}), independently confirming the model
  card's rtol = 1e-4. The governance conclusions carry to kUPS's JAX
  runtime exactly.
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\section{One-paragraph summary}\label{one-paragraph-summary}

On the locked positive control, audit found the recipe had
\emph{double-counted silicon} --- Talu's Ar--O cross-parameter already folds
silicon's dispersion into oxygen, and the recipe added an explicit Ar--Si
on top --- so its strict-band agreement was a fortuitous four-error cancellation; rebuilt clean
(single-source oxygen-only, converged convention, on Talu's calibration
Olson geometry) it reproduces the corrected reference (\(K_H = 0.200\))
exactly and cross-engine (native 0.200, RASPA3 0.199). Separately and
intact, the engines agree on the shared recipe to 1.1 \% --- the
cross-engine parity result. CuspAI's own
Widom estimator was independently validated for the first time --- its
energies are right (q\_st to 0.22 kJ/mol) and its +16 \% K\_H was
root-caused to a shifted-vs-truncated LJ convention (δ ≈ 0.43 kJ/mol,
reconciling both K\_H and q\_st to within the SEM). Pointed at the MLIP
CuspAI ships, governance discriminated two distinct failure modes
(overlap over-binding vs missing dispersion), screen-passed a
D3(BJ)-corrected route on the single governed Si-CHA + Ar screen, and
showed the deployment-class flagship (UMA) fails the same way
--- with the governance harness proving its own validity for each engine
before judging it. \emph{The published table is never the whole recipe;
the unpublished conventions are load-bearing, and they can be
characterized, discriminated, and in some cases fixed.}
